Handling String-Number Conversion in String Constraint Solving

### 2019-11-19 draft.tex
update implementation and evaluation section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-pfa}}}

@@

\section{Implementation}\label{section:implementation}
\input{evaluation}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-PFA}}}

@@

\section{Implementation}\label{section:implementation}
The tool {\tool} is done by implementing our string constraint solving procedure on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enable us to solve not only formulas over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulas.

In {\tool}, we use the following PFA selection strategy: we select PFA's of size $m$ for string variables occur a number-string conversion constraint and PFA's with $p$ loops of size $q$ for other variables. Initially, we set $(m,p,q)=(X,X,X)$, and increase each of them by one if more precise refinement is required. We set an upper bound $7$ for each parameters and report UNKNOWN if a solution cannot be found within the bound. \yfc{need to check with Diep for the parameters}.

The way our theory solver and Z3 interacts is almost standard. Every time when Z3 asks our theory solver a problem in string theory, our theory solver tries to prove it UNSAT or SAT using the aforementioned procedure. For under-approximation, every time when a linear formula for under-approximation is created, we create a new solver for linear integer arithmetic inside Z3 to solve it. If our theory solver reports UNKNOWN, Z3 remembers it in a global flag \textsf{incomplete} and tries another solution branch. If all Z3 completed the search of all solution branches, it reports UNSAT if the flag \textsf{incomplete} is down, and UNKNOWN otherwise.

\input{evaluation}

### 2019-11-19 evaluation.tex
update implementation and evaluation section

\todo{Add comparison/discussion; References of benchmarks; Update data; Sepatate categories of str\_int benchmark for discussion}

@@

We have implemented the proposed approach in a prototype tool \tool as a part of the Z3 theorem prover. This approach allows us to efficiently solve not only formulae over string constraints, but also combine string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

We compare \tool with other state-of-the-art string solvers, namely CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, \textsf{z3str3}~\cite{zheng2013z3}, Ostrich~\cite{chen2019decision} and \textsf{Trau+}~\cite{parosh2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints. We present data on two benchmark groups that demonstrate two points. First, only our tool can efficiently solve formulae that contain string-number and number-string conversion constraints and which can be a combination of constraints over multiple theories. Second, we show that our tool is competitive on conversion-less benchmarks as well. The summary of experimenting for each benchmark group is given in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. All experiments were executed on a computer with 4-core CPU, 8 GiB RAM and Ubuntu 18.4-LVM OS. The timeout was set to 10s for each test. Columns with heading \texttt{sat}/\texttt{unsat} show numbers of solved formulae. Columns with heading \texttt{timeout} indicate the number of times the solver exceeded the time limit and columns with heading $\times$ show numbers of instances for which the particular solver either reported unknown or reported incorrect answer or crashed.

Occasionally, during the experiments, there was a case where a solver reported a different answer to a test than the other solvers. To decide which solver reported the incorrect answer, we have developed our own validator. Our validator takes the model from the solver who reported that the test is satisfiable and assigns the corresponding values according to the returned model to the variables in the test. The modified test is re-evaluated by solvers. If the solvers decide that the modified test is satisfiable, the answer of the solvers that initially reported unsat is evaluated as incorrect.

>>>>

@@

We compare {\tool }with other state-of-the-art string solvers, namely CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, \textsf{z3str3}~\cite{zheng2013z3}, Ostrich~\cite{chen2019decision} and \textsf{Trau+}~\cite{parosh2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints, which occurs in most of our benchmarks. We present data on two benchmark groups that demonstrate two points. First, {\tool} is the most efficient tool for formulas that contain string-number conversion constraints. Second, {\tool} is among the best tools in on benchmarks without string-number conversion constraints.
The results of the experiments is summarized in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. All experiments were executed on a computer with 4-core CPU, 8 GiB RAM and Ubuntu 18.4-LVM OS. The timeout was set to 10s for each test. Columns with heading \texttt{sat}/\texttt{unsat} show numbers of solved formulas. Columns with heading $\times$ indicate the number of instances the solver fails (including timeout, report unknown, report incorrect result, or crashed).

Occasionally, during the experiments, there was a case where a solver reported a different answer to a test than the other solvers. To decide which solver reported the incorrect answer, we have developed our own validator. Our validator takes the model from the solver who reported that the test is satisfiable and assigns the corresponding values according to the returned model to the variables in the test. The modified test is re-evaluated by multiple solvers. If all solvers decide that the modified test is satisfiable, the answer of the solvers that initially reported unsat is evaluated as incorrect.
